---
title: "Some Programming in Base R"
author: "Cass Crews"
format: html
editor: visual
---

# Task 1

We are analyzing blood pressure before and after a medical treatment. 

### 1.

First, we will create the vectors corresponding to pre-treatment and post-treatment blood pressures. 

```{r}
#Pre-treatment blood pressure
pre_treat_bp<-c(120, 151, 125, 126, 115, 132, 132, 129, 134, 139, 127, 122, 
                127, 135, 133, 128, 147, 138, 140, 132)

#Post-treatment blood pressure
post_treat_bp<-c(127, 145, 135, 122, 115, 122, 123, 126, 126, 129, 132, 146, 
                 120, 114, 121, 120, 128, 120, 133, 115)
```


### 2.

Next, we will assign names to each observation in each vector, corresponding to the subject each observation belongs to. 

```{r}
#Assigning names to pre-treatment vector
names(pre_treat_bp)<-paste("Subject", 1:20, sep = "_")

#Assigning names to post-treatment vector
names(post_treat_bp)<-paste("Subject", 1:20, sep = "_")
```

### 3.

Next, we will take the difference between pre-treatment blood pressures and post-treatment pressures. The subject-specific differences are printed below. 

```{r}
#Calculating differences in BP
diff_bp<-pre_treat_bp-post_treat_bp

#Printing the differences
diff_bp
```

### 4. 

Next, we will calculate the average difference across the 20 subjects: 5.65.  

```{r}
#Calculating average difference
mean_diff_bp<-mean(diff_bp)

#Printing average
mean_diff_bp
```

### 5.

Next, we will identify the subjects that experienced a decrease in blood pressure following treatment. Those 15 subjects are listed below. 

```{r}
#Identifying which subjects experienced a decrease in BP
which(diff_bp> 0)
```

### 6.

Next, we will subset the vector of differences to capture only those subjects that experienced a decrease in blood pressure (positive change). The resultant vector is listed below. 

```{r}
#Creating a vector of only the positive differences
pos_diff_bp<-diff_bp[diff_bp> 0]

#Printing vector
pos_diff_bp
```

### 7. 

Finally, we will calculate the average decrease among only those that experienced a decrease: 10.6

```{r}
#Calculating average difference for only positive changes
mean_pos_diff_bp<-mean(pos_diff_bp)

#Printing average
mean_pos_diff_bp
```

